A Convolutional Neural Network for Outsole Recognition\tnoteref{t1}

### 2019-06-10 paper/paper.Rnw
Add and restructure consistency images, update outline of results, and other miscellaneous updates

\mt{Ultimately, most pattern evidence is stored in photographic form; as a result, we must use image analysis methods to create numerical features from the original visual record.} There are a number of methods that may be employed to identify shapes and features in an image, such as Fourier transforms, Hough transformations \citep{ballardGeneralizingHoughTransform1981} and other low-level feature extraction methods aimed at detecting specific shapes, such as edges, corners, blobs, or ridges \citep[Ch 15]{machineVision}. While these methods are useful in identifying these specific features at a low level, they can be computationally intensive and only identify features on a very small scale; as a result, they cannot reliably identify large geometric shapes like those that may be found in an outsole image.

%\svp{Motivate the use of CNNs directly - complex model, higher level of feature aggregation. Mention pre-trained networks to transition.}
Convolutional neural networks (CNNs) are widely recognized as superior for novel image classification. CNNs are a form of artificial neural network which make use of the image convolution operator used by many low-level feature extraction methods, with the additional ability to aggregate such features and meaningfully connect them to a pre-determined set of labels. CNNs have deep architectures that can be trained to identify complex patterns, but they are structurally similar to the architecture of the human visual system and output binary or probabilistic predictions for given labels that are readily interpretable. As CNNs make use of labeled training data, the predictions generated are for features which are similar to those identified by humans, resulting in models with greater face validity. Once a CNN is trained, it is relatively fast and easy to apply the model to new images and obtain classifications.

\paragraph{Transfer Learning}
Neural networks are composed of sets of layers; the early layers, which are called the model base, contain feature detectors, and the final layers make up the classifier, or model head, which connects meaningful features to classification labels. A relatively simple convolutional neural network which processes image data may contain more than 14 million parameters in the model base and an additional 120 million parameters in the model head; to train a model of this nature requires millions of labeled images and a significant amount of computational power. Assembling data to train a network from scratch is often a nearly impossible task; the process of \emph{transfer learning} is a natural solution to this common problem. Transfer learning leverages the modularity of neural networks, that is, that the model base can be separated from the classifier which produces predictions.

Many convolutional neural networks have been trained on a set of images known as ImageNet; these images are hierarchically labeled, so the same image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity \citep{dengImageNetLargeScaleHierarchical}. CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously. As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks.
\svp{You need to talk about the model structure of vgg16 before you start talking about layers. Talk of convolutional blocks doesn't really make sense until you show the layer structure.}
\mt{I agree, but I thought that this was where we were arguing that the ``eyes" are transferrable, so I wanted to talk about the filters to back that up.}
\autoref{fig:vgg16layers} shows a selection of filters from VGG16, where each filter detects the presence of specific shapes and colors. Each row of filters in \autoref{fig:vgg16layers} corresponds to one of the five convolutional blocks of VGG16. Filters in earlier convolutional blocks detect simple features, such as colors and blobs, while later filters detect more complex combinations of features. Each of the five convolutional blocks has between 128 and 1,536 filters, and a single image is made up of some combination of the thousands of features that are detected by those filters.

@@

\fix{In this paper, we will use transfer learning to automatically identify features in shoe treads which are used by forensic examiners, leveraging a general-purpose CNN with relatively simple structure, VGG16, and a database of labeled shoe tread images we have assembled for this project, which are both described below.} \svp{This sentence is a bit awkward, because you're introducing too many things. The paragraph is good, but I don't think that it makes sense to have it here. Introduce VGG16, then use this paragraph as a transition to talking about the actual model we fit.} This approach differs from other approaches \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017} in automatic footwear identification which use the output from the model base directly and do not attempt to add human-friendly contextual information with an additional classifier.

@@

\paragraph{VGG16}
Developed by Oxford's Visual Graphics Group, VGG16 is a CNN with 16 ``functional" (i.e., convolutional and densely connected) layers and 5 ``structural" max pooling layers \citep{krizhevskyImageNetClassificationDeep2012}. In contrast to other popular networks, like ResNet, VGG has a relatively simple structure that provides easier training and interpretability with very little sacrificed accuracy. The simplicity of this structure provides the ability to peer into the inner workings of the network for diagnostic purposes, providing a distinct advantage over more complicated network structures with slightly higher accuracy ratings. VGG16 is a common choice for transfer learning because of this structural simplicity; it has been used for detection of text in natural images \citep{zhongDeepTextUnifiedFramework2016}, medical imaging classification \citep{oquabLearningTransferringMidlevel2014}, classification of weld defects \citep{liuWeldDefectImages2018}, and many other domain-specific image recognition tasks that are more specific than the ImageNet data on which it was trained. \autoref{fig:VGG16-structure} shows the architecture of VGG16; with transfer learning, the VGG16 model head is replaced with a trained model head specific to recognition of shoeprint class characteristics.

@@

library(magick)
b <- image_blank(width = 300, height = 1590)
f <- image_read(file.path(wd, "images", "unlabeled_filters.png"))
f_l <- image_append(c(b, f)) %>%
  image_annotate("Block 1:", location = "+60+159",
                 color = "black", size = 60) %>%
  image_annotate("Block 2:", location = "+60+477",
                 color = "black", size = 60) %>%
  image_annotate("Block 3:", location = "+60+795",
                 color = "black", size = 60) %>%
  image_annotate("Block 4:", location = "+60+1113",
                 color = "black", size = 60) %>%
  image_annotate("Block 5:", location = "+60+1431",
                 color = "black", size = 60) %>%
  image_scale(geometry = "600x")
image_write(image = f_l, path = file.path(wd, "images", "labeled_filters.png"),
            format = "png")
knitr::include_graphics(file.path(wd, "images", "labeled_filters.png"))
@

In order to train a new head for the VGG16 model base, human-labeled images are necessary. The next section describes how these images were produced and used to fit our custom classifier.

@@

The categories we use in this study are operationally defined as follows:
\begin{description}
\item [Bowtie] Bowtie shapes are roughly quadrilateral, with two opposite concave faces. The remaining two faces can be convex or straight, and the concave faces may have straight portions, so long as there is a concave region. %Using this definition, shapes such as butterflies are included as bowties.
\item [Chevron] Chevron shapes include repeating parallel lines as well as individual ``v" shapes. They may be angular but can also be curved.
\item [Circle] Circles include ellipses and ovals; they must be round.
\item [Line] Lines are repeated and parallel; a more general definition of a line would be difficult to differentiate from many other patterns. Lines can be mildly curved.
\item [Polygon] Polygons are defined in this standard to have more than 4 sides. They include pentagons, hexagons, and octagons.
\item [Quadrilateral] Quadrilaterals (quads) have four sides. They may have rounded or square corners.
\item [Star] Stars are any shape with alternating concave and convex regions, or lines which emanate from a central point. ``X" and ``+" shapes are also classified as stars.
\item [Text] Text is any shape which would be identified as text by a reasonable human. In most cases, the text on the outsole images used is made up of Latin alphabet characters; the model will likely not recognize text in other scripts (but could be trained if non-Latin text images could be obtained).
\item [Triangle] Triangles are any three-sided figure. Like quadrilaterals, they can have rounded corners. In some cases, it is difficult to distinguish between a trapezoidal shape and a triangle when rounded corners are involved.
\item [Other] Other features which were marked include logos, various textures (including crepe, stippling, etc.), and smooth regions with no discernible features. These regions are grouped and provide additional information - that none of the previous nine categories are present. \mt{Why no EM dash? ``---"}
\end{description}

@@

<<class-characteristic-barchart, fig.width = 6, fig.height = 4, dpi = 300, fig.cap = "Distribution of classes in all labeled images. Quadrilaterals, lines, circles, text, and chevrons are relatively common; stars, polygons, and bowties are relatively rare.", fig.scap = "Distribution of classes in all labeled images.">>=
annotated_imgs %>%
  mutate(label_type = str_replace(label_type, "multi", "multiple")) %>%
  mutate(labels = str_to_title(labels)) %>%
  mutate(labels = factor(labels, levels = c("Quad", "Line", "Text", "Circle",
                                            "Chevron", "Triangle", "Polygon",
                                            "Star", "Bowtie", "Other"))) %>%
  ggplot() +
  geom_bar(aes(x = labels, fill = label_type), color = "black") +
  scale_fill_manual("Labels", values = c("single" = "#6ba2b9",
                                         "multiple" = "#2e5597")) +
  coord_flip() +
  ylab("# Labeled Images") +
  xlab("")  +
  ggtitle("Current Class Distribution (All Labeled Images)") +
  mytheme +
  theme(legend.position = c(1, 1),
        legend.justification = c(1.03, 1.03),
        legend.background = element_rect(fill = "white"))
@

%Labeled images are scarce relative to the amount of data necessary to train a neural network.
Although transfer learning substantially reduces the amount of data required to train a CNN by several orders of magnitude, \fix{labeled images are still difficult or time-consuming to generate relative to the amount of data necessary to train even a small neural network adequately. Another common solution to data scarcity is to artificially enlarge the data set using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}.}
\svp{I would lose the last clause of the first sentence and combine it with the next, e.g. to make most efficient use of the labeled data, we enlarge the training data using a process called image augmentation.}
\mt{I don't think the first half of the first sentence is relevant here without the second. Should I use a different transition? I was trying to put augmentation in the context of transfer learning; i.e., transfer learning helps, but getting data is hard, so we use augmentation to get more, with the added benefit of improving generalizability.}
Augmentation is the transformation of original input data using image operations such as cropping, zoom, skew, rotation, and color balance modification in order to distort or alter the image while maintaining the essential features corresponding to the label. This process reduces the potential for overfitting the model to the specific set of image data used during the training process, and also increases the amount of data available for training. Examples of pre- and post-augmentation images are shown in \autoref{fig:augmented}.

@@

\fix{How much should be said about loss and loss functions?}\svp{}
A loss function is a function that measures how far model predictions are from the given labels for each image in the validation and test sets; the model ``learns" by updating its predictions to minimize the loss.

\autoref{fig:training-accuracy} shows the training and validation accuracy and loss at each epoch of the fitting process. Overfitting occurs when a model learns the training data so well that its understanding of the categories becomes specific to the training cases, which, in turn, leads to poor prediction of new images that were not in the training set.
%Overfitting, or fitting a model which performs too well on the training data relative to the validation data,
Overfitting occurs when the validation loss begins to increase after reaching a global minimum. Alternately, underfitting occurs if the validation accuracy is still increasing when model optimization is terminated (for instance, between epochs 1 and 10 in \autoref{fig:training-accuracy}), because model performance is still improving with continued training. In \autoref{fig:training-accuracy}, validation accuracy levels off after epoch 20, and validation loss has not yet begun to increase at epoch 30, indicating that the model optimization process was halted at an appropriate epoch.

@@

\mt{\textbf{OUTLINE OF RESULTS}
\begin{itemize}
\item ROC shows that, in general, model performance is better than chance and broadly compares class performance.
\item Show model predictions primarily to provide a few examples to back up the claim that predictions generally look good, but also include one or two case studies that motivate class-by-class diagnostics
\item Confusion matrix shows which classes are most accurate or most problematic. However, when we see that one shape is commonly misclasified as another, the conf mat doesn't let us do anything more than speculate the cause.
\item Heatmaps help us look at specific instances of classification to 1) verify that the model is using appropriate features when it classifies correctly and 2) diagnose the cause of confusion when the model classifies incorrectly
\item (Potentially switch the order of prediction examples and the confusion matrix, so the conf mat indicates systemic class confusion and then back that up with the examples?)
\end{itemize}
}

\paragraph{Overall Accuracy} % Model-wise and class-by-class, w/ confusion matrix
%\svp{You need to explain these terms. Your CC was written for statisticians, this paper is for practitioners. You need a MUCH gentler introduction to the idea of accuracy, errors, and ROC curves, as well as a lot more definitions for terms.}

For each image in the test set, the ``true labels" are nine human-assigned labels of 0 or 1 corresponding to \svp{the presence or absence of} each of the nine shape categories \svp{in the image}. When the CNN predicts which shapes are in the image, however, it assigns a probability between 0 and 1 for each shape category, with a total of nine probabilities per test image. Thus, to determine the accuracy of the model predictions, we must select a threshold for each label to discretize the model predictions. %such that a predicted probability above the threshold indicates a ``yes" prediction from the model and a probability below the threshold indicates a ``no".}
If the threshold is too high, \fix{moderate} probabilities will be treated as a ``no", which increases the number of true negative and false negative predictions. Conversely, if the threshold is too low, \fix{moderate} probabilites will be considered as ``yes", which increases the true and false positive predictions in the data. An appropriate threshold must be chosen to produce a high true positive rate while also ideally keeping the false positive rate low.

%\mt{I don't immediately see how to separate the below sentence into two without making it really clunky... I figured that in the context of thresholds (previous paragraph), it would be easy to just stick in that those thresholds are used to compute the curve.}

%Better?: Receiver Operating Characteristic (ROC) curves are a type diagnostic plot which compare the false positive rate to the true positive rate for a classification method at a number of thresholds, and the Area Under the Curve (AUC) quantifies the shape of the ROC curve.

Receiver Operating Characteristic (ROC) curves are a type diagnostic plot which compare the false positive rate to the true positive rate for a classification method.\svp{Make this two sentences so that it's less convoluted}\fix{, computed by comparing the two rates at a number of thresholds, and the Area Under the Curve (AUC) quantifies the shape of the ROC curve.} Perfect prediction would be indicated by a right angle along the upper left corner of the plot, with a corresponding AUC of 1, and a diagonal line with an AUC of 0.5 would indicate that the classification method performs no better than random chance. \autoref{fig:overall-roc} shows the ROC curve for \fix{the model} across all classes, and \autoref{fig:class-roc} shows the curve for each class. The full model has an AUC of 0.88, and the AUC for individual classes ranges from 0.81 (for lines) to 0.91 (for bowtie and text). While the class performances do vary slightly, each ROC curve is the same general shape and performs significantly better than random chance. The points in \autoref{fig:class-roc} represent the equal error rate (EER), which is the threshold where there is equal probability of a false positive or false negative error. We will use the class-wise EER to set a threshold corresponding to each class, so that any model prediction with probability above the EER for that shape is considered a positive prediction.

@@

\svp{Pros and cons of confusion matrix here? Maybe we just don't show the actual proportions on the graph? I'd like to discuss this more in person.}

\svp{I'd focus more on the fact that ROC curves are a general summary of the model's performance, but we must also examine the specific cases where the model misclassifies objects to determine if there is a pattern.}
\mt{ROC curves provide a general summary of model accuracy, but they do not indicate which classes are being assigned the false positive or false negative labels when accuracy is not perfect. A confusion matrix is a helpful diagnostic tool to understand which classes are commonly confused for other classes.} A confusion matrix is a \mt{summary of the relationships between} the labels assigned to an image and the labels predicted by the model.

In the confusion matrix presented in \autoref{fig:confusion_matrix}, the values along the diagonal represent the proportion of true positives captured within each category. The off-diagonal values represent the proportion of false positives for each shape pairing; however, since a single image may truly contain multiple shapes, the values have been adjusted to remove the effect of any true positives from the calculation of false positive proportions. \svp{As a result of this adjustment, the probabilities off-diagonal are the proportion ... need to explain the conditional probability we're showing here} %\svp{I can't remember exactly how we're doing that, but we may want to look more closely at that - if we set everything to NA instead of 0, the rate calculations would be higher than expected} \mt{They're being set to NA, and then we're calculating proportions with na.rm = T. I don't understand what you mean about higher rates than expected...}.
For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any image that truly contains both triangles and quadrilaterals is removed before calculating the proportion of false quadrilateral labels.

@@

\paragraph{Model Predictions}

 \mt{What else to say here?}

<<exemplars-consistency, fig.cap = "Examples", fig.pos = "p!", out.width = "100%", cache = T, fig.pos = "!h">>=
exemplar <- list.files(file.path(wd, "images", "consistency", "exemplar"), full.names = T)
pred_prob_plot(exemplar, loaded_model)
@

\fix{Can also look at many examples within same shape categories.}

<<chevrons-consistency, fig.cap = "Chevrons", fig.pos = "p!", out.width = "100%", cache = T>>=
chevron <- list.files(file.path(wd, "images", "consistency", "chevron"), full.names = T)
pred_prob_plot(chevron, loaded_model)
@

<<circle-text-consistency, fig.cap = "Text with circles or circle-like shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
circle_text <- list.files(file.path(wd, "images", "consistency", "circle-text"), full.names = T)
pred_prob_plot(circle_text, loaded_model)
@

\svp{I've moved this down here because I think it fits better in the model exploration section - using this to explain why performance differs - intraclass variability.}

>>>>

Ultimately, most pattern evidence is stored in photographic form; as a result, we must use image analysis methods to create numerical features from the original visual record. There are a number of methods that may be employed to identify shapes and features in an image, such as Fourier transforms, Hough transformations \citep{ballardGeneralizingHoughTransform1981} and other low-level feature extraction methods aimed at detecting specific shapes, such as edges, corners, blobs, or ridges \citep[Ch 15]{machineVision}. While these methods are useful in identifying these specific features at a low level, they can be computationally intensive and only identify features on a very small scale; as a result, they cannot reliably identify large geometric shapes like those that may be found in an outsole image.

Convolutional neural networks (CNNs) are widely recognized as superior for novel image classification. CNNs are a form of artificial neural network which make use of the image convolution operator used by many low-level feature extraction methods, with the additional ability to aggregate such features and meaningfully connect them to a pre-determined set of labels. CNNs have deep architectures that can be trained to identify complex patterns, but they are structurally similar to the architecture of the human visual system and output binary or probabilistic predictions for given labels that are readily interpretable. As CNNs make use of labeled training data, the predictions generated are for features which are similar to those identified by humans, resulting in models with greater face validity. Once a CNN is trained, it is relatively fast and easy to apply the model to new images and obtain classifications.

\paragraph{Transfer Learning}
%\mt{Training a CNN from scratch is hard. There are CNNs trained on complex images, and transfer learning lets us use those weights. In particular, we're using VGG16. We can look at the filters of VGG16 to support our claim that it will be good for the task.}
Neural networks are composed of sets of layers; the early layers, which are called the model base, contain feature detectors, and the final layers make up the classifier, or model head, which connects meaningful features to classification labels. A relatively simple convolutional neural network which processes image data may contain more than 14 million parameters in the model base and an additional 120 million parameters in the model head; to train a model of this nature requires millions of labeled images and a significant amount of computational power. Assembling data to train a network from scratch is often a nearly impossible task; the process of \emph{transfer learning} is a natural solution to this common problem. Transfer learning leverages the modularity of neural networks, that is, that the model base can be separated from the classifier which produces predictions.

Many convolutional neural networks have been trained on a \mt{sub}set of images \mt{from the database} ImageNet; these images are hierarchically labeled, so the same image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity \citep{dengImageNetLargeScaleHierarchical}. CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously. As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks. \fix{More to be said later.}

@@

@@

\paragraph{VGG16}
Developed by Oxford's Visual Graphics Group, VGG16 is a \mt{pre-trained} CNN with 16 ``functional" (i.e., convolutional and densely connected) layers and 5 ``structural" max pooling layers \citep{krizhevskyImageNetClassificationDeep2012}. In contrast to other popular networks, like ResNet, VGG has a relatively simple structure that provides easier training and interpretability with very little sacrificed accuracy. The simplicity of this structure provides the ability to peer into the inner workings of the network for diagnostic purposes, providing a distinct advantage over more complicated network structures with slightly higher accuracy ratings. VGG16 is a common choice for transfer learning because of this structural simplicity; it has been used for detection of text in natural images \citep{zhongDeepTextUnifiedFramework2016}, medical imaging classification \citep{oquabLearningTransferringMidlevel2014}, classification of weld defects \citep{liuWeldDefectImages2018}, and many other domain-specific image recognition tasks that are more specific than the ImageNet data on which it was trained. \autoref{fig:VGG16-structure} shows the architecture of VGG16; with transfer learning, the VGG16 model head is replaced with a trained model head specific to recognition of shoeprint class characteristics.

\mt{The model base of VGG16 is made up of five convolutional blocks alternated with max pooling layers. Each convolutional block is made up of two or three convolutional layers, and each layer is made up of hundreds of filters, such that each of the five convolutional blocks has between 128 and 1,536 filters. Strictly speaking, a filter is a set of numerical weights that, when applied to an image, quantifies the presence of the specific shapes and/or colors that the filter is trained to detect. \autoref{fig:vgg16layers} shows a selection of filters from VGG16, with each row corresponding to one of the five convolutional blocks. Filters in earlier convolutional blocks detect simple features, such as colors and blobs, while later filters detect more complex combinations of features. When considered together, the series of filters perform the process of feature detection, such that a single image is made up of some combination of the thousands of features that are detected by filters. Although the filters of VGG16 have been trained and optimized on the ImageNet Data, the large number of filters span a wide variety of shapes and features} \fix{which are still useful/relevant to classifying outsoles. In a sense, this is much like teaching humans to classify objects they have never seen before: an object may be unfamiliar, but our eyes can still detect its color, shape, and texture because we have learned these ideas from other sources.}

@@

library(magick)
b <- image_blank(width = 300, height = 1590)
f <- image_read(file.path(wd, "images", "unlabeled_filters.png"))
f_l <- image_append(c(b, f)) %>%
  image_annotate("Block 1:", location = "+60+139",
                 color = "black", size = 60) %>%
  image_annotate("Block 2:", location = "+60+457",
                 color = "black", size = 60) %>%
  image_annotate("Block 3:", location = "+60+775",
                 color = "black", size = 60) %>%
  image_annotate("Block 4:", location = "+60+1093",
                 color = "black", size = 60) %>%
  image_annotate("Block 5:", location = "+60+1411",
                 color = "black", size = 60) %>%
  image_scale(geometry = "600x")
image_write(image = f_l, path = file.path(wd, "images", "labeled_filters.png"),
            format = "png")
knitr::include_graphics(file.path(wd, "images", "labeled_filters.png"))
@

In this paper, we demonstrate the use of transfer learning to automatically identify features in shoe treads that are used by forensic examiners. To do this, we leverage the relatively simple structure and generalizability of VGG16 and train a new classifier using a database of labeled shoe tread images we have assembled for this project. This approach differs from other approaches \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017} in automatic footwear identification that use the output from the model base directly and do not attempt to add human-friendly contextual information with an additional classifier. The next section describes how the database of images was produced and used to fit our custom classifier.

@@

The categories we use in this study are operationally defined as follows:
\begin{description}
\item [Bowtie] Bowtie shapes are roughly quadrilateral, with two opposite concave faces. The remaining two faces can be convex or straight, and the concave faces may have straight portions, so long as there is a concave region. %Using this definition, shapes such as butterflies are included as bowties.
\item [Chevron] Chevron shapes include repeating parallel lines as well as individual ``v" shapes. They may be angular but can also be curved.
\item [Circle] Circles include ellipses and ovals; they must be round.
\item [Line] Lines are repeated and parallel; a more general definition of a line would be difficult to differentiate from many other patterns. Lines can be mildly curved.
\item [Polygon] Polygons are defined in this standard to have more than 4 sides. They include pentagons, hexagons, and octagons.
\item [Quadrilateral] Quadrilaterals (quads) have four sides. They may have rounded or square corners.
\item [Star] Stars are any shape with alternating concave and convex regions, or lines which emanate from a central point. ``X" and ``+" shapes are also classified as stars.
\item [Text] Text is any shape which would be identified as text by a reasonable human. In most cases, the text on the outsole images used is made up of Latin alphabet characters; the model will likely not recognize text in other scripts (but could be trained if non-Latin text images could be obtained).
\item [Triangle] Triangles are any three-sided figure. Like quadrilaterals, they can have rounded corners. In some cases, it is difficult to distinguish between a trapezoidal shape and a triangle when rounded corners are involved.
\item [Other] Other features which were marked include logos, various textures (including crepe, stippling, etc.), and smooth regions with no discernible features. These regions are grouped and provide the additional information that none of the previous nine categories are present.
\end{description}

@@

<<class-characteristic-barchart, fig.width = 6, fig.height = 4, dpi = 300, out.width = "80%", fig.cap = "Distribution of classes in all labeled images. Quadrilaterals, lines, circles, text, and chevrons are relatively common; stars, polygons, and bowties are relatively rare.", fig.scap = "Distribution of classes in all labeled images.">>=
annotated_imgs %>%
  mutate(label_type = str_replace(label_type, "multi", "multiple")) %>%
  mutate(labels = str_to_title(labels)) %>%
  mutate(labels = factor(labels, levels = c("Quad", "Line", "Text", "Circle",
                                            "Chevron", "Triangle", "Polygon",
                                            "Star", "Bowtie", "Other"))) %>%
  ggplot() +
  geom_bar(aes(x = labels, fill = label_type), color = "black") +
  scale_fill_manual("Labels", values = c("single" = "#6ba2b9",
                                         "multiple" = "#2e5597")) +
  coord_flip() +
  ylab("# Labeled Images") +
  xlab("")  +
  ggtitle("Current Class Distribution (All Labeled Images)") +
  mytheme +
  theme(legend.position = c(1, 1),
        legend.justification = c(1.03, 1.03),
        legend.background = element_rect(fill = "white"))
@

%Labeled images are scarce relative to the amount of data necessary to train a neural network.
Although transfer learning substantially reduces the amount of data required to train a CNN by several orders of magnitude, labeled images are still \fix{difficult or time-consuming} to generate relative to the amount of data necessary to train even a small neural network adequately. Thus, to make the most efficient use of the existing labeled data, we enlarge the training data using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}. Augmentation is the transformation of original input data using image operations such as cropping, zoom, skew, rotation, and color balance modification in order to distort or alter the image while maintaining the essential features corresponding to the label. This process reduces the potential for overfitting the model to the specific set of image data used during the training process, and also increases the amount of data available for training. Examples of pre- and post-augmentation images are shown in \autoref{fig:augmented}.

@@

In machine learning, a loss function is a function that measures how far model predictions are from the given labels for each image in the validation and test sets; the model ``learns" by updating its predictions to minimize the loss. \autoref{fig:training-accuracy} shows the training and validation accuracy and loss at each epoch of the fitting process.

Overfitting is when a model learns the training data so well that its understanding of the categories becomes specific to the training cases, which, in turn, leads to poor prediction of new images that were not in the training set. In training, overfitting is indicated when the validation loss begins to increase after reaching a global minimum. Alternately, underfitting occurs if the validation accuracy is still increasing when model optimization is terminated (for instance, between epochs 1 and 10 in \autoref{fig:training-accuracy}), because model performance is still improving with continued training. In \autoref{fig:training-accuracy}, validation accuracy levels off after epoch 20, and validation loss has not yet begun to increase at epoch 30, indicating that the model optimization process was halted at an appropriate epoch.

@@

\mt{\textbf{OUTLINE OF RESULTS}
\begin{itemize}
\item ROC
  \begin{itemize}
  \item Overall: model performance is better than chance
  \item Classes: broad comparison of class performance
  \end{itemize}
\item Examples
  \begin{itemize}
  \item Show what predictions look like
  \item Introduce the idea that some classifications aren't perfect
  \item Motivate systematic class-by-class confusion
  \end{itemize}
\item Confusion Matrix
  \begin{itemize}
  \item Show patterns in misclassification
  \item Conf mat doesn't tell us why certain predictions are wrong
  \end{itemize}
\item In this paper, we identify two main reasons for error: some classes are more variable, and some classes share physical features (I.e., the model has to learn a lot of features to encapsulate a class with lots of variety, but then other classes share those features)
\item Brands
  \begin{itemize}
  \item Bowties/polygons/stars are mostly a few brands, but the top quad/line brands still only take up a small proportion of all quads (.04 or less)
  \end{itemize}
\item Heatmaps
  \begin{itemize}
  \item Verify that the model is using appropriate features when predictions are correct
  \item Show examples for strong misclassification, but make the case that the mistakes are understandable
  \end{itemize}
\end{itemize}
}

\paragraph{Overall Accuracy}

For each image in the test set, the ``true labels" are nine human-assigned labels of 0 or 1 corresponding to the presence or absence of each of the nine shape categories in the image. When the CNN predicts which shapes are in the image, however, it assigns a probability between 0 and 1 for each shape category, with a total of nine probabilities per test image. Thus, to determine the accuracy of the model predictions, we must select a threshold for each label to discretize the model predictions.
If the threshold is too high, \fix{moderate} probabilities will be treated as a ``no", which increases the number of true negative and false negative predictions. Conversely, if the threshold is too low, \fix{moderate} probabilites will be considered as ``yes", which increases the true and false positive predictions in the data. An appropriate threshold must be chosen to produce a high true positive rate while also ideally keeping the false positive rate low.

Receiver Operating Characteristic (ROC) curves are a type diagnostic plot which compare the false positive rate to the true positive rate for a classification method, and the Area Under the Curve (AUC) quantifies the shape of the ROC curve. Perfect prediction would be indicated by a right angle along the upper left corner of the plot, with a corresponding AUC of 1, and a diagonal line with an AUC of 0.5 would indicate that the classification method performs no better than random chance. \autoref{fig:overall-roc} shows the ROC curve for \fix{the model} across all classes, and \autoref{fig:class-roc} shows the curve for each class. The full model has an AUC of 0.88, and the AUC for individual classes ranges from 0.81 (for lines) to 0.91 (for bowtie and text). While the class performances do vary slightly, each ROC curve is the same general shape and performs significantly better than random chance. The points in \autoref{fig:class-roc} represent the equal error rate (EER), which is the threshold where there is equal probability of a false positive or false negative error. We will use the class-wise EER to set a threshold corresponding to each class, so that any model prediction with probability above the EER for that shape is considered a positive prediction.

@@

\paragraph{Model Predictions}
\mt{ROC curves provide a general summary of model accuracy. To learn more about how specific images are classified, we can view images from the test set along with their respective nine model predictions. \autoref{fig:single-exemplars} shows model predictions for images that contain only one type of shape, and \autoref{fig:multi-exemplars} shows similar predictions for images containing multiple shapes. Most predictions for the single-label images are straightforward, but there is much less consensus of model opinion in the multi-label images. }

<<single-exemplars, fig.cap = "Examples containing a single shape", fig.pos = "p!", out.width = "100%", cache = T>>=
archetype <- list.files(file.path(wd, "images", "consistency", "single-exemplar"), full.names = T)
pred_prob_plot(archetype, loaded_model, sort = F)
@

<<multi-exemplars, fig.cap = "Examples containing multiple shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
exemplar <- list.files(file.path(wd, "images", "consistency", "multi-exemplar"), full.names = T)
pred_prob_plot(exemplar, loaded_model)
@

\mt{These figures provide representative examples of classifications across multiple categories; it is also useful to examine many cases within a given shape. \autoref{fig:chevrons-consistency}}

<<chevrons-consistency, fig.cap = "Chevrons", fig.pos = "p!", out.width = "100%", cache = T>>=
chevron <- list.files(file.path(wd, "images", "consistency", "chevron"), full.names = T)
pred_prob_plot(rev(chevron), loaded_model, sort = F)
@

<<circle-text-consistency, fig.cap = "Text with circles or circle-like shapes", fig.pos = "p!", out.width = "100%", cache = T>>=
circle_text <- list.files(file.path(wd, "images", "consistency", "circle-text"), full.names = T)
pred_prob_plot(rev(circle_text), loaded_model, sort = F)
@

%\svp{Pros and cons of confusion matrix here? Maybe we just don't show the actual proportions on the graph? I'd like to discuss this more in person.}

\fix{Now that we've seen specific examples, it is useful to summarize the systematic errors by class.} A confusion matrix is a helpful diagnostic tool to understand which classes are commonly confused for other classes.} A confusion matrix is a \mt{summary of the relationships between} the labels assigned to an image and the labels predicted by the model.

In the confusion matrix presented in \autoref{fig:confusion_matrix}, the values along the diagonal represent the proportion of true positives captured within each category. The off-diagonal values represent the proportion of false positives for each shape pairing; however, since a single image may truly contain multiple shapes, the values have been adjusted to remove the effect of any true positives from the calculation of false positive proportions. \svp{As a result of this adjustment, the probabilities off-diagonal are the proportion ... need to explain the conditional probability we're showing here} %\svp{I can't remember exactly how we're doing that, but we may want to look more closely at that - if we set everything to NA instead of 0, the rate calculations would be higher than expected}
For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any image that truly contains both triangles and quadrilaterals is removed before calculating the proportion of false quadrilateral labels.

@@
